---
title: "Week 02:"
subtitle: "Describing Data in R"
author: "Paul Testa"
output:
  xaringan::moon_reader:
    css: ["default", "css/brown.css"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-lakeside-light
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
  comment = NA, dpi = 300,
  fig.align = "center", out.width = "80%", cache = FALSE)
library("tidyverse")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


---
class: inverse, center, middle
# Overview

---
## General Plan

- Review
- Lecture consists of a mix of:
  - Topics and concepts
  - Exercises
- Summary

---
## Review


---
## Topics

- Working with R Markdown files

- Basic Programming in R

- Setting up your work space

- Loading data into R

- Looking at your data

- Cleaning and transforming your data

- Describing what's typical of your data

- Describing how your data vary

---
## Exercises

1. Create an R Markdown File

2. Explore objects and functions in base `R`

3. Install the COVID19 package

4. Load data on COVID19 

5. Inspect the data

6. Welcome to the tidyverse

7. Measures of central tendency

8. Measures of dispersion

---
class: inverse, center, middle

# Key Concept

## Comment


---
## Point 1


---
## Code Slide

.left-column[
```{r,echo=T,eval=F}
2+2
```


]

.right-column[
```{r,eval=T}
plot(1:10,1:10)
```

]

---
class: inverse, center, middle

# Key Concept
