---
title: "Week 02:"
subtitle: "Casual Infrence in Experimental Designs"
author: "Paul Testa"
output:
  xaringan::moon_reader:
    css: ["default", "css/brown.css"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-lakeside-light
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
  comment = NA, dpi = 300,
  fig.align = "center", out.width = "80%", cache = FALSE)
library("tidyverse")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)
```



class: inverse, center, middle
# Overview

---
## General Plan

- Review
- Lecture consists of a mix of:
  - Topics and concepts
  - Exercises
- Summary

---
## Review

- Data wrangling
- Data visualization

---
## Data transformations


You want to do:

- 
- 
- 
- 
- 
- 


---
## Data visualization

A basic graphic requires at minimum:

- `data:` the dataset containing the variables of interest.
- `aes`: aesthetic attributes of the geometric object. For example, x/y position, color, shape, and size. Aesthetic attributes are mapped to variables in the dataset.
- `geom:` the geometric object in question. This refers to the type of object we can observe in a plot For example: points, lines, and bars.
---
## Data visualization

Basic graphics are made even better with:

- `facets`
- `statistics`
- `coordinates`
- `themes`




---
# ðŸ”
# Recoding categorical data in R

---
## Topics

- Causal inference

- Causal Identification

- A potential outcomes framework causal inference

- A statistical solution to the fundamental problem causal inference

- How to estimate causal effects 


---
## Exercises

1. What kinds questions have causal interpretations?

2. Estimating an average treatment effect with the `resume` data

3. Exploring the data for Broockman and Kalla 2016



---
class: inverse, center, middle
# ðŸ’¡
# Causal Inference
## *Causal claims imply counterfactual comparisons*

---
## Casual claims are are all around us

.pull-left[

]

.pull-right[

]

---
## Casual claims are are all around us





---
## Causal claims imply counterfactual comparisons


---
class: inverse, center, middle
# ðŸ’¡
# Causal Identification
## *What assumptions do we need to make for a causal claim to be credible*


---
## Observational vs Experimental Designs

---
class: inverse, center, middle
# ðŸ’¡
# A Potential Outcomes Framework for Causal Inference
## The fundamental problem of causal inference is that, for an individual unit, we only observe one of potentially many potential outcomes $\to$ Individual causal effects are unidentified

---
# Potential Outcomes


---
class: inverse, center, middle
# ðŸ’¡
# The fundamental problem of causal inference

---
class: inverse, center, middle
# ðŸ’¡
# A statistical solution to the fundamental problem

---
class: inverse, center, middle
# ðŸ’¡
# Randomization as a reasoned basis for inference


---
class: center, middle
# ðŸ’ª  
# What kinds questions have causal interpretations?

---
class: center, middle
# ðŸ’ª  
# Estimating an average treatment effect with the `resume` data

```{r}
library(qss)
library(tidyverse)
data("resume")



```

```{r}
dim(resume)
head(resume)
```

```{r}
resume %>%
  group_by(sex, race) %>%
  summarize(
    Y = mean(call)
  )

resume %>%
  group_by(sex) %>%
  summarize(
    Y = mean(call)
  )

resume %>%
  filter(!firstname %in% c("Kristen","Carrie","Brad","Jay"))%>%
  group_by(sex) %>%
  summarize(
    Y =  mean(call[race =="white"]) - mean(call[race =="black"])
  )

resume %>%
  group_by(race) %>%
  summarize(
    Y = mean(call)
  )

resume %>%
  group_by(race, sex,firstname)%>%
  summarize(
    Y = mean(call),
    n = n()
  )%>%
  arrange(desc(Y)) %>%
  mutate(
    firstname = forcats::fct_reorder(firstname,Y)
  )%>%
  ggplot(aes(Y, firstname,col=race, size=n))+
  geom_point() + 
  facet_grid(sex~.,scales = "free_y")
```


```{r}

```






---
class: center, middle
# ðŸ’ª  
# Exploring the data for Broockman and Kalla 2016


