---
title: "Week 08:"
subtitle: "Probably too much Probability Theory"
author: "Paul Testa"
output:
  xaringan::moon_reader:
    css: ["default", "css/brown.css"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-lakeside-light
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
  comment = NA, dpi = 300,
  fig.align = "center", out.width = "80%", cache = FALSE)
library("tidyverse")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)
```



```{r packages, include=F}
the_packages <- c(
  ## R Markdown
  "kableExtra","DT","texreg", "htmltools",
  ## Tidyverse
  "tidyverse", "lubridate", "forcats", "haven", "labelled",
  ## Extensions for ggplot
  "ggmap","ggrepel", "ggridges", "ggthemes", "ggpubr", 
  "GGally", "scales", "dagitty", "ggdag", "ggforce",
  # Graphics:
  "scatterplot3d", #<<
  # Data 
  "COVID19","maps","mapdata","qss","tidycensus", "dataverse", 
  # Analysis
  "DeclareDesign", "zoo", "gtools"
)
```

```{r ipak, include=F}
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}

```


```{r loadpackages, cache=F, include=F}
ipak(the_packages)
```

---
class: inverse, center, middle
# Overview

---
## General Plan

- Setup
- Feedback
- Review
- Lecture 
  - Topic 1
  - Topic 2

- Summary


---
class:inverse, middle, center
# 💪
## Get set up to work

---
## New packages

Hopefully, you were all able to install the following packages 

```{r, eval=FALSE}


```



---
## Packages for today


```{r, ref.label=c("packages")}

```

---
## Define a function to load (and if needed install) packages


```{r, ref.label="ipak"}
```

---
## Load packages for today

```{r ref.label="loadpackages"}
```


---
class:inverse, center, middle
# 💪
## Load Data for today


# 📢
## Feedback


```{r feedback, echo=F, message=F}
df_feedback_full <- haven::read_spss("../files/data/wk05.sav")
df_feedback_full %>%
  filter(optout == 1)-> df_feedback
```


---
## What we liked

--

- Likes here

---
## What we liked

```{r likes, echo=F, out.height='90%'}
DT::datatable(df_feedback %>% 
                select(Likes),
               fillContainer = F,
              height = "90%",
              options = list(
                pageLength = 4
              )
              )
```


---
## What we disliked

--

- Dislikes here
---
## What we disliked

```{r dislikes, echo=F}
DT::datatable(df_feedback %>% 
                select(Dislikes),
               fillContainer = F,
              height = "90%",
              options = list(
                pageLength = 3
              )
              )
```


---
class:inverse, middle, center
# 🔍
## Review

---
## Predicted Values


---
## Overfitting vs Under Fitting




---
class: inverse, center, middle
# 💡
# WARNING

---
## Where we're headed.

---
## What you should learn from today

- Language for randomness

- Probability is tool for quantifying and describing randomess

- Random variables

- Probability Distributions

- Expected values and variance 

---
class: inverse, center, middle
# 💡
# Randomness

## Random Experiments, Sets Theory, and Counting Rules


---
## Random Experiments and Sample Spaces

- In probability theory, a **Random Experiment:** describes a repeatable process where the outcome is uncertain (non-deterministic)

- The **Sample Space** is the **set** of all possible outcomes of an experiment ($S$) 


**Examples**

- Flip a coin 

- Flip a coin 10 times and count the number of heads

- Number of wars in the 20th century

- Outcome of an election

---
## Sets

- In mathematics, a set is bounded collection of elements

- If $A$ is the set of outcomes from a coin toss then $A=\{Heads,Tails\}$

---
## Elements

- An object within a set

- `1` is an element in set of outcomes from rolling a six-sided die 

---
## Events

- An Event is a collection of outcomes of an experiment

  - Could contain a single element, multiple elements, no elements, or all the elements.

- If the event $B$ is rolling an odd number, then $B=\{1,3,5\}$

- The outcome `3` is an element of $B$ ($3 \in B$)

- The outcome `4` is not an element of $B$ ($4 \notin B$)


---
## How many elements in a set?

A set can be:

- **Countably Finite:** 
  
  - Roll a die, there six possible outcomes $\{1,2,3,4,5,6\}$


- **Countably Infinite:** 

  - Number of rolls until a six appears ${1,2,3,\dots}$

--

- **Uncountably Infinite:** 

  - All the real numbers between 0 and 1 $(0,1)$


---
## How many possible outfits could you have made me wear?

---
## How to count


The Fundamental Counting Principle says that if there $x$ ways to do one thing and $z$ ways to do another then, then are their are $x \times z$ total ways of doing both tasks.

More generally, if there are 

- If there are $j$ tasks or decision stages

- And $k$ choices at each decision stage $j$ such that $n_{j,k} = k$

- The total number of possible outcomes is the product of the number number of choices $k$ and each stage, $j$ $\prod_{i=1}^{j} n_k = n_{i,k}*n_{2,k}*\dots*n_{j,k}$ 


---
## How many different outfits could you have made me wear?

- **Palette**: $\{\text{Fall, Winter, Spring, Summer}\}$
- **Jacket**: $\{\text{Tuxedo, Tweed,Blazer, Sportcoat, No coat}\}$
- **Top**: $\{\text{Dress Shirt, Polo, Tee Shirt, Sports Jersey}\}$
- **Pant**: $\{\text{Slacks, Khakis, Jeans, Shorts}\}$
- **Shoe**: $\{\text{Dress Shoe, Dress boot, Jordans, Dunks, Basketball, Crocs}\}$
- **Tie**: $\{\text{Repp, Pattern, Knit, Bow, No tie}\}$
- **Pattern**: $\{\text{Simple, Stripes, Checks, Graphic}\}$

- How many decision stages?

- How many choices at each stage?

- How many possible outfits?

---
## How many different suggested outfits could you have made me wear?

- How many decision stages? $j=7$

- How many choices at each stage? 

- **Palette**: $\{\text{Fall, Winter, Spring, Summer}\} = n_{jk} = 4 $ 
- **Jacket**: $\{\text{Tuxedo, Tweed,Blazer, Sportcoat, No coat}\} = n_{jk} = 5$
- **Top**: $\{\text{Dress Shirt, Polo, Tee Shirt, Sports Jersey}\} = n_{jk} = 4$
- **Pant**: $\{\text{Slacks, Khakis, Jeans, Shorts}\} = n_{jk} = 4$
- **Shoe**: $\{\text{Dress Shoe, Dress boot, Jordans, Dunks, Basketball, Crocs}\} = n_{jk} = 6$
- **Tie**: $\{\text{Repp, Pattern, Knit, Bowtie, No tie}\} = n_{jk} = 5$
- **Pattern**: $\{\text{Simple, Stripes, Checks, Graphic}\} = n_{jk} = 4$

- How many possible suggestions? $\prod_{i=1}^{j} n_{jk} = 4 \times 5 \times 4 \times 4 \times 6 \times 5 \times 4 = 38,400 \text{ outfits}$

---
## And yet you chose this:

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">I send students in my undergraduate research class a weekly survey to get a sense of what&#39;s working and what&#39;s not. <br><br>I also ask them silly questions just for fun. Last week, I asked them to help me with my fit. This was the result: <a href="https://t.co/w1t8Cr04Ap">pic.twitter.com/w1t8Cr04Ap</a></p>&mdash; Paul Testa (@ProfPaulTesta) <a href="https://twitter.com/ProfPaulTesta/status/1501291241404833795?ref_src=twsrc%5Etfw">March 8, 2022</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


---
## Permutations and Combinations

- A permutation is a selection of objects in which the order matters.

  - There are *six* ways to arrange 


---
## Counting with Regard to Order and Replacement

In general, when we think about the total number of outcomes, it matters whether 

- we are counting *with* or *without* replacement
- we can distinguish the order of observations



```{r, echo=F}
counting_tab <- tibble::tibble(
   ` `=  c("Permutation", "Perumation","Combination", "Combination"),
  `Order Matters` = c("Yes","Yes","No","No"),
  `Repetition Allowed` = c("Yes","No","No","Yes"),
  Formula = c("$P(n,k) = {}_{n}P_k = n^k $", 
              "$P(n,k) = {}_{n}P_k = \\frac{n!}{(n-k)!}$",
              "$C(n,k) = {}_{n}C_k = \\binom{n}{k} = \\frac{{}_{n}P_k}{k!} \\frac{n!}{k!(n-k)!}$",
              "$C(n,k) = {}_{n}C_k = \\binom{n + k - 1}{k} =\\frac{(n + k -1)!}{k!(n-1)!}$"
              ),
  "Example = {abc}" = c('P(3,3) = Pick three letters in order with replacement = $3^3 = 27$ permutations',
              'P(3,3) = Pick three letters in order without replacement = $\\frac{3!}{(3-3)!} = \\frac{3*2*1}{1} = 6$ permutations',
              'C(3,3) = Choose three letters without replacement = $\\frac{3!}{3!(3-3)!} = \\frac{3*2*1}{3*2*1} = 1$ combination',
              'C(3,3) = Choose three letters with replacement = $\\frac{3+3 -1!}{3!(3-1)!} = \\frac{5*4*3*2*1}{(3*2*1)(2*1)} = 10$ combinations'
              )
  )


choose(3,3)
knitr::kable(counting_tab)%>%kableExtra::kable_styling()
```

---

```{r}
permutations(3, 2, v = c("a","b","c"),repeats.allowed = T)
permutations(3, 2, v = c("a","b","c"),repeats.allowed = F)
```

```{r}
combinations(3, 2, v = c("a","b","c"),repeats.allowed = F)  
combinations(3, 2, v = c("a","b","c"),repeats.allowed = T)

```


---
## Example

- Roll a die twice, record the outcome. How many possible outcomes are there?\pause
- 36 ($n^k=6^2=36$)

## Sample Space

```{r,echo=F,results='asis'}
two.die<-matrix(NA,6,6)
for(i in 1:6){
    for(j in 1:6)
        two.die[i,j]<-paste("{",i,", ",j,"}",sep="")
}
rownames(two.die)<-paste("\\textbf{",1:6,"}")
colnames(two.die)<-paste("\\textbf{",1:6,"}")
library(xtable)
print(xtable(two.die),sanitize.text.function = force,comment = F)
```

---
## Permutations

- An urn with six tickets numbered 1 through 6.
    - Draw 1 ticket, record number, eat ticket (e.g. w/o replacement)
    - Draw another ticket, record number.
    - How many unique pairs of numbers can you get?

---    
## Permutations

- Counting with regard to order and without replacement \pause   
- **Answer** 
\[
\begin{aligned}
\textrm{P}^6_2&=\frac{n!}{(n-k)!}\\
            &=\frac{6!}{(6-2)!} \\
            &=\frac{6*5*4*3*2*1}{4*3*2*1}\\
            &=\frac{6*5}{1}=30
\end{aligned}
\]

---
## Permutations

```{r,echo=F,results='asis'}
two.die.p<-two.die
for(i in 1:6){
two.die.p[i,i]<-""
}

print(xtable(two.die.p),sanitize.text.function = force,comment = F)
```

---
## Combinations

- Counting without regard to order and without replacement
- How many unique combinations of tickets are their?

- **Answer** 
\[
\begin{aligned}
\textrm{C}^6_2&=\binom{n}{k}\\
            &=\frac{6!}{2!(6-2)!} \\
            &=\frac{6*5*4*3*2*1}{(4*3*2*1)(2*1)}\\
            &=\frac{6*5}{2*1}=15
\end{aligned}
\]

---
## Combinations

```{r,echo=F,results='asis'}
two.die.c<-two.die
two.die.c[lower.tri(two.die.c,diag=T)]<-""

print(xtable(two.die.c),sanitize.text.function = force,comment = F)
```

---
## Without regard to order, with replacement

- Roll two die, how many pairings can we observe?

```{r,echo=F,results='asis'}
two.die.craps<-two.die
two.die.craps[lower.tri(two.die.craps,diag=F)]<-""

print(xtable(two.die.craps),sanitize.text.function = force,comment = F)
```



---
## Operations on Sets

- Empty Set: $\o$ a set with no elements
- Subset: 
    - $D=\{1,2,3,4,5,6\}$
    - $Primes=\{2,3,5\}$
    - $Primes \subset D \iff \forall X \in Primes, X \in D$
- Unions
    - $A \cup B = \{X:X \in A \lor X \in B \}$
    - Either $A$, $B$ or both $A and B$ occur
- Intersections
    - $A \cap B = \{X:X \in A \land X \in B \}$
    - Both $A$ and $B$ occur
- Complements
    - $A^\complement = \{X:X\notin A\}$
    - $A^\complement$ means $A$ does not occur
    - $\o^\complement=S$ and $S^\complement=\o$ 

---
## Union

\begin{center}
\begin{tikzpicture}
\draw (-2,-1.5) rectangle (3.5,1.5) node[below left]{$S$};
\fill[gray] (0,0) circle (1cm);
\fill[gray] (1.5,0) circle (1cm);
\draw (0,0) circle (1cm) node {$A$};
\draw (1.5,0) circle (1cm) node {$B$};
\end{tikzpicture}
\end{center}

---
## Intersection

\begin{center}
\begin{tikzpicture}
\draw (-2,-1.5) rectangle (3.5,1.5) node[below left]{$S$};
\begin{scope} % start of clip scope
\clip (0,0) circle (1cm);
\fill[gray] (1.5,0) circle (1cm);
\end{scope} % end of clip scope
\draw (0,0) circle (1cm) node[left] {$A$};
\draw (1.5,0) circle (1cm) node[right] {$B$};
\end{tikzpicture}
\end{center}

---
## Complement of Intersection

\begin{center}
\begin{tikzpicture}
\filldraw[gray] (-2,-1.5) rectangle (3.5,1.5) node[below left]{$S$};
\begin{scope} % start of clip scope
\clip (0,0) circle (1cm);
\fill[white] (1.5,0) circle (1cm);
\end{scope} % end of clip scope
\draw (0,0) circle (1cm) node[left] {$A$};
\draw (1.5,0) circle (1cm) node[right] {$B$};
\end{tikzpicture}
\end{center}

---
## Properties of Sets

### Commutative
$A \cup B = B \cup A$
$A \cap B = B \cap A$


### Associative
$A\cap (B \cap C) = (A \cap B)\cap C$
$A\cup (B \cup C) = (A \cup B)\cup C$

---
## Properties of Sets

### Distributive
$A\cap (B \cup C)=(A\cap B)\cup (A \cap C)$
$A\cup (B \cap C)=(A\cup B)\cap (A \cup C)$

### de Morgan's Law
- $(A \cup B)^\complement = A^\complement \cap B^\complement$
- $(A \cap B)^\complement = A^\complement \cup B^\complement$


---
## Mutually Exclusive (Pairwise Disjoint)

- $A_i\cap A_j = \o \forall i \neq j$
- $\bigcup_{i=1}^k A_i = S \rightarrow A_1 \dots A_k$ is a **partition**  of the sample space


---
class: inverse, center, middle
# 💡
# Probability


---
class: inverse, center, middle
# 💡
# Conditional Probability




---
class: inverse, center, middle
# 💡
# Random Variables and Probability Distributions

---
class: inverse, center, middle
# 💡
# Expected Values and Variances


---
class: inverse, center, middle
# 💡
# Common Probability Distributions





