---
title: "Week 10:"
subtitle: "Quantifying uncertainty with confidence intervals"
author: "Paul Testa"
output:
  xaringan::moon_reader:
    css: ["default", "css/brown.css"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-lakeside-light
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
  comment = NA, dpi = 300,
  fig.align = "center", out.width = "80%", cache = FALSE)
library("tidyverse")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)
```



```{r packages, include=F}
the_packages <- c(
  ## R Markdown
  "kableExtra","DT","texreg",
  ## Tidyverse
  "tidyverse", "lubridate", "forcats", "haven", "labelled",
  ## Extensions for ggplot
  "ggmap","ggrepel", "ggridges", "ggthemes", "ggpubr", 
  "GGally", "scales", "dagitty", "ggdag", "ggforce",
  # Graphics:
  "scatterplot3d", #<<
  # Data 
  "COVID19","maps","mapdata","qss","tidycensus", "dataverse", 
  # Analysis
  "DeclareDesign", "easystats", "zoo"
)
```

```{r ipak, include=F}
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}

```


```{r loadpackages, cache=F, include=F}
ipak(the_packages)
```

---
class: inverse, center, middle
# Overview

---
## General Plan

- Setup
- Feedback
- Review
- Lecture 
  - Topic 1
  - Topic 2

- Summary


---
class:inverse, middle, center
# 💪
## Get set up to work

---
## New packages

Hopefully, you were all able to install the following packages 

```{r, eval=FALSE}


```



---
## Packages for today


```{r, ref.label=c("packages")}

```

---
## Define a function to load (and if needed install) packages


```{r, ref.label="ipak"}
```

---
## Load packages for today

```{r ref.label="loadpackages"}
```


---
class:inverse, center, middle
# 💪
## Load Data for today

We'll use data from last weeks lab to 

```{r}
load(url("https://pols1600.paultesta.org/files/data/df_drww.rda"))
```



# 📢
## Feedback


```{r feedback, echo=F, message=F}
df_feedback_full <- haven::read_spss("../files/data/wk05.sav")
df_feedback_full %>%
  filter(optout == 1)-> df_feedback
```


---
## What we liked

--

- We liked the lab (or liked it better)

- Working in groups 

- Extensive notes

- Reasonable pacing 

- Real world data

- Mix of substantive and technical questions

---
## What we liked

```{r likes, echo=F, out.height='90%'}
DT::datatable(df_feedback %>% 
                select(Likes),
               fillContainer = F,
              height = "90%",
              options = list(
                pageLength = 4
              )
              )
```


---
## What we disliked

--

- Labs still feel rushed (at the end)

- When code breaks it can be hard to fix

- Lecture was a bit confusing/long

- The space

---
## What we disliked

```{r dislikes, echo=F}
DT::datatable(df_feedback %>% 
                select(Dislikes),
               fillContainer = F,
              height = "90%",
              options = list(
                pageLength = 3
              )
              )
```


---
class:inverse, middle, center
# 🔍
## Review: Generalized Linear Models


---
## Topics

- Working with R Markdown files

- Basic Programming in R

- Setting up your work space

- Loading data into R

- Looking at your data

- Cleaning and transforming your data

- Describing what's typical of your data

- Describing how your data vary

---
## Exercises

1. Create an R Markdown File

2. Explore objects and functions in base `R`

3. Install the COVID19 package

4. Load data on COVID19 

5. Inspect the data

6. Welcome to the tidyverse

7. Measures of central tendency

8. Measures of dispersion

---
class: inverse, center, middle
# 💡
# Confidence Intervals
## The Basics


---
## Overview:





---
## Defintions: Populations and Samples


- **Population**: All the cases from which you could have sampled

- **Parameter:** A quantity or quantities of interest often generically called $\theta$ ("theta"). Something we'd like to know about our population

- **Sample:** A (random) draw from that population

- **Sample Size:** The number of observations in your draw (without replacement)


---
## Defintions: Estimators, Estimates, and Statistics

- **Estimator:** A rule for calculating an *estimate* of our parameter of interest. 

- **Estimate:** The value produced by some estimator for some parameter from some data. Often called $\hat{\theta}$ 

- **Unbiased estimators:** $E(\hat{\theta})=E(\theta)$ On average, the estimates produced by some estimator will be centered around the truth

- **Consistent estimates:** $\lim_{n\to \infty} \hat{\theta_N} = \theta$ As the sample size increases, the estimates from an estimator converge in probability to the parameter value

- **Statistic:** A summary of the data (mean, regression coefficient, $R^2$). An estimator without a specified target of inference 

---
## Definitions: Distrubtions and Standard Errors

- **Sampling Distribution:** How some estimate would vary if you took repeated samples from the population

- **Standard Error:** The standard deviation of the sampling distribution

- **Resampling Distribution:** How some estimate would vary if you took repeated samples **from your sample WITH REPLACEMENT** 
    - "Sampling from our sample, as the sample was sampled from the population."
    
---
## Confidence Intervals

- Confidence intervals give a range of values that are likely to include the true value of the parameter $\theta$ with probability $(1-\alpha) \times 100\%$

  - $\alpha = 0.05$ corresponds to a "95-percent confidence interval"

---
## Confidence Intervals: Interpretation

- Our "confidence" is about the interval
  
- In repeated sampling, we expect that $(1-\alpha) \times 100\%$ of the intervals we construct would contain the truth.

- For any one interval, the truth, $\theta$, either falls within in the lower and upper bounds of the interval or it does not.

---
## Two Approaches to Calculating Confidence Intervals:

In general, there are two ways to calculate confidence intervals:

- **Simulation:** Use our computers to simulate the idea of repeated sampling (e.g. bootstrapping)

  - Flexible, but more computationally intensive

- **Asymptotic Theory:** Use math to derive the properties of the distributions that would arise under repeated sampling
  
  - Faster, but requires more assumptions that may not hold

We will consider both. 

- The theory of CIs is easier to illustrate via simulation

- The practice of calculating CIs is (generally) easier using asymptotic theory



---
class: inverse, center, middle
# 💡
# Confidence Intervals
## Simulating the Sampling Distribution through Bootstrapping


---
## Populations

Let's load the data from the *Do Russians Want War* survey

```{r}
load(url("https://pols1600.paultesta.org/files/data/df_drww.rda"))

```

To understand the logic of confidence intervals, let's treat this data as our **population** from which we we could draw repeated samples. 

---
## Population Age


In our population, there are **parameters**, true values of things we want to know. 

Suppose we're interested in the average age of our population.

In our population, the true value of $\mu_{age} = E[Age]$ is

```{r}
mu_age <- mean(df_drww$age)
mu_age
```

Similarly, the true $\sigma_{age} = \sqrt{E[Age^2] - E[Age]^2}$

```{r}
sd_age <- sqrt(mean((df_drww$age-mean(df_drww$age))^2))
sd_age
```


---
## Distribution Population Age


```{r}
p_pop <- df_drww %>%
  ggplot(aes(age))+
  geom_density(col="grey")+
  geom_rug()+
  geom_vline(
    aes(xintercept = mu_age, 
             col = "Population Mean"),
    linetype=2)+
  theme_bw()+
  labs(color = "Age")

p_pop
```


---
## Sample Estimates of Average Age (N = 25)

Suppose we took three samples, **without replacement** of size 25, and calculated the average age in each sample:

```{r}
set.seed(123)
mean_age1 <- mean(sample(df_drww$age, 25, replace = F))
mean_age2 <- mean(sample(df_drww$age, 25, replace = F))
mean_age3 <- mean(sample(df_drww$age, 25, replace = F))

mean_age1
mean_age2
mean_age3
```


---
```{r}
df_mn <- tibble(
  xint = c(mu_age, mean_age1, mean_age2, mean_age3),
  sample = c("Population","Sample 1", "Sample 2","Sample 3"),
  distribution = c("Sampling","Bootstrap", "Bootstrap","Bootstrap")
)

p_samp <- df_drww %>%
  ggplot(aes(age))+
  geom_density(col="grey")+
  geom_rug()+
  geom_vline(
    data = df_mn,
    aes(xintercept = xint, 
             col = sample,
        linetype=sample)
    )+
  theme_bw()+
  labs(color = "Age")
p_samp

```

---
## Repeated Sampling

- Imagine we could draw a 1,000 or 10,000 or an infinite number of samples of size N=25 from our population.

- How much would our estimate of the average of age of the population vary?

- Let's use our computers to simulate this process and find out!

---
## Simualting Repeated Sampling

```{r}
library(boot)
library(purrr)
n_sims <- 1000
samp_size <- 25
set.seed(123)

mu_age_samp_dist_n25 <- tibble(
  sim = 1:n_sims,
  distribution = "Sampling",
  sample = "Population"
) %>%
  mutate(
    samp = map(sim, ~ slice_sample(df_drww, n = samp_size, replace = F)),
    estimate = map_dbl(samp, ~ mean(.$age))
  )

```

---

```{r}
mu_age_samp_dist_n25

```

---
# One Sample

```{r}
mu_age_samp_dist_n25$samp[[1]]$age
mean(mu_age_samp_dist_n25$samp[[1]]$age)
mu_age_samp_dist_n25$estimate[[1]]
```


```{r, echo=F}
mu_age_samp_dist_n25 %>%
  group_by(sim)%>%
  group_map(.,~bind_rows(.$samp)) -> tmp
class(tmp)

df_sims <- bind_rows(mu_age_samp_dist_n25$samp)
df_sims%>%
  mutate(
    num_id = sort(rep(rep(1:1000),25)),
    id = forcats::fct_reorder(paste("Sample ", sort(rep(rep(1:1000),25))),num_id)
  ) -> df_sims

df_sims_samp <- df_sims %>%
  dplyr::group_by(id)%>%
  dplyr::summarise(
    num_id = unique(num_id),
    mean = mean(age)
  )

df_sims %>%
  filter(num_id <13)%>%
ggplot(aes(age, group = id))+
  geom_density()+
  geom_rug()+
  facet_wrap(~id)+
  geom_vline(
    data = df_sims_samp%>%filter(num_id <13),
    aes(xintercept = mean, group = id),
    linetype = 2
  )+
  labs(title="Distribution of Age in 12 Samples")+
  theme_bw()


```


---

```{r}
p_dist <- mu_age_samp_dist_n25 %>%
  ggplot(aes(estimate))+
  geom_density()+
  geom_rug()+
  geom_density(
    data = df_drww,
    aes(x=age),col = "grey"
  )+
   geom_vline(
    aes(xintercept = mu_age, 
             col = "Population Mean"),
    linetype=2)+
  theme_bw()+
  labs(title = "Distribution of Sample Means (N=25)")
p_dist
```


---
## Standard Errors

- A **standard error** is simply the standard deviation of the sampling distribution.

- The standard error for our simulation above:

```{r}
se_age_n25 <- sd(mu_age_samp_dist_n25$estimate)
se_age_n25
```

---
## Coverage Intervals

- From the Central Limit Theorem, we know that the distribution of sample means will converge to a normal distribution.

- From probability theory, we know that we that roughly 95 percent of the values in a normal distribution fall between *Two* Standard Deviations of the mean.

```{r}
ci_age_ul_n25 <- mu_age + 2*se_age_n25
ci_age_ll_n25 <- mu_age - 2*se_age_n25

mean(mu_age_samp_dist_n25$estimate >ci_age_ll_n25 & 
       mu_age_samp_dist_n25$estimate <ci_age_ul_n25)

```


---
```{r}
mu_age_samp_dist_n25 %>%
  ggplot(aes(estimate))+
  geom_density()+
  geom_rug(
    aes(col = estimate >ci_age_ll_n25 & 
          estimate <ci_age_ul_n25)
  )+
  geom_vline(xintercept = mu_age,
             col = "red",
             linetype=2)+
  guides(col="none")+
  geom_segment(aes(x=ci_age_ll_n25,
                   xend = ci_age_ul_n25,
                   y = .15,yend = .15 ),
               col = "#00BFC4")+
  theme_bw()
```

---
## Boostrapped Standard Errors 

- A standard error is the standard deviation of a hypothetical sampling distribution

- How do we calculate a standard error from a single sample?

- It turns out that a random sample provides unbiased estimates of both the population mean **and** the standard deviation of the of the sampling distribution (i.e. the standard error).

- We can estimate this this standard error, by sampling with replacement from our sample to generate a **bootstrapped** sampling distribution


---
## Boostrapped Standard Errors

```{r}
set.seed(123)
bs_resamp_1 <- tibble(
  sim = 1:nsims,
  distribution = "Bootstrap",
  sample = "Sample 1",
) %>%
  mutate(
    samp = map(sim, ~ slice_sample(
      mu_age_samp_dist_n30$samp[[1]], n = samp_size, replace = T)),
    estimate = map_dbl(samp, ~ mean(.$age))
  )
```


---
## Boostrapped Standard Errors

```{r}
bs_resamp_2 <- tibble(
  sim = 1:nsims,
  distribution = "Bootstrap",
  sample = "Sample 2",
) %>%
  mutate(
    samp = map(sim, ~ slice_sample(
      mu_age_samp_dist_n30$samp[[2]], n = samp_size, replace = T)),
    estimate = map_dbl(samp, ~ mean(.$age))
  )
```

---
## Boostrapped Standard Errors

```{r}
bs_resamp_3 <- tibble(
  sim = 1:nsims,
  distribution = "Bootstrap",
  sample = "Sample 3",
) %>%
  mutate(
    samp = map(sim, ~ slice_sample(
      mu_age_samp_dist_n30$samp[[3]], n = samp_size, replace = T)),
    estimate = map_dbl(samp, ~ mean(.$age))
  )
```


---
## Boostrapped Standard Errors

```{r}
bs_example <- rbind(
  mu_age_samp_dist_n25,
  bs_resamp_1,
  bs_resamp_2,
  bs_resamp_3
)

df_se <- bs_example %>%
  select(sample, estimate)%>%
  dplyr::group_by(sample)%>%
  dplyr::summarise(
    se = sd(estimate)
  )


df_ci <- df_mn %>%
  left_join(df_se)

df_ci%>%
  mutate(
    ll = xint - qt(df=25,.975)*se,
    ul = xint + qt(df=25,.975)*se,
    y = .15,
    xint_pop = xint[1]
  ) -> df_ci

```

---
## Boostrapped Standard Errors
```{r}
bs_example %>%
  ggplot(aes(estimate,col = sample))+
  geom_density(aes(linetype=distribution))+
  facet_wrap(~sample, ncol=1)+
   geom_vline(
    data = df_ci,
    aes(xintercept = xint, 
             col = sample,
        linetype=distribution)
    )


bs_example %>%
  ggplot(aes(estimate,col = sample))+
  geom_density(aes(linetype=distribution))+
  facet_wrap(~sample, ncol=1)+
   geom_vline(
    data = df_ci,
    aes(xintercept = xint, 
             col = sample,
        linetype=distribution)
    )+
    geom_vline(
    data = df_ci,
    aes(xintercept = xint_pop), 
             col = "black",
        linetype=2)+
  geom_segment(
    data = df_ci,
    aes(x =ll, xend =ul, y=y, yend=y)
  )


```

---
## Simulating Confidence Intervals for Sample Sizes

```{r}
sim_ci_fn<-function(x,
                    samp_size=100,
                    n_sims=1000,
                    level=.95,
                    bs=F){
    # Take a sample of size "nsamp"
    y<-sample(x=na.omit(x),size=samp_size,replace=F)
    # Calculate the mean
    mu<-mean(y,na.rm=T)
    # If bs=TRUE do bootstrapped SEs 
    if(bs==T){
    mu_dist<-rerun(
      n_sims,
      mean(sample(y, samp_size, replace = T)))%>%
      unlist()
    se<-sd(mu_dist)}else{
    # Otherwise, just use assymptotic result (Quicker)
    se<-sd(y,na.rm=T)/sqrt(samp_size-1)
    }
    # Significance level
    the.p<-1-(1-level)/2
    # Calculate lower and upper limits of interval
    ll<-mu-qt(p=the.p,df=samp_size-1)*se
    ul<-mu+qt(p=the.p,df=samp_size-1)*se
    results<-tibble(mu=mu,ll=ll,ul=ul,se=se)
    return(results)
    
}
set.seed(12345)
samp25 %>%
  dplyr::mutate(
    sample = 1:n()
  )
samp25 <- map_df(1:1000, ~sim_ci_fn(df_drww$age, samp_size = 25)) %>%dplyr::mutate(sample = 1:n() )
samp50 <- map_df(1:1000, ~sim_ci_fn(df_drww$age, samp_size = 50))%>%dplyr::mutate(sample = 1:n() )
samp100 <- map_df(1:1000, ~sim_ci_fn(df_drww$age, samp_size = 100))%>%dplyr::mutate(sample = 1:n() )
samp200 <- map_df(1:1000, ~sim_ci_fn(df_drww$age, samp_size = 200))%>%dplyr::mutate(sample = 1:n() )

samp1000$ul
samp500 %>%
  summarise(
    coverage = mean(ll<mu_age&ul>mu_age)
  )
samp25 %>%
  summarise(
    coverage = mean(ll<mu_age&ul>mu_age)
  )
samp50 %>%
  summarise(
    coverage = mean(ll<mu_age&ul>mu_age)
  )
samp100 %>%
  summarise(
    coverage = mean(ll<mu_age&ul>mu_age)
  )
samp200 %>%
  summarise(
    coverage = mean(ll<mu_age&ul>mu_age)
  )

# Coverage

test <- samp200 %>%
  slice_sample(n=100) %>%
  ggplot(
    aes(x=mu,
        y=sample,
        xmin=ll,
        xmax=ul,
        col=ll<mu_age&ul>mu_age))+
  geom_point()+
  geom_vline(xintercept = mu_age)+
  geom_errorbarh()+
  scale_color_discrete(guide="none")+
  labs(title="95% CIs for 200 samples of\nSize 10")
test
new <- samp25%>%slice(1:100)
test %+% samp25[1:100,]
test %+% samp50[1:100,]
test %+% samp100[1:100,]
test %+% samp200[1:100,]
ggplot(samp500,aes(x=mu,y=sample,xmin=ll,xmax=ul,col=ll<mu_age&ul>mu_age))+geom_point()+geom_vline(xintercept = mu_age)+
    geom_errorbarh()+scale_color_discrete(guide=F)+labs(title="95% CIs for 100 samples of\nSize 10")
```










  

