---
title: "Week XX:"
subtitle: "Topic HERE"
author: "Paul Testa"
output:
  xaringan::moon_reader:
    css: ["default", "css/brown.css"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-lakeside-light
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
  comment = NA, dpi = 300,
  fig.align = "center", out.width = "80%", cache = FALSE)
library("tidyverse")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
  ),
  rmarkdown::html_dependency_font_awesome()
)
```



```{r packages, include=F}
the_packages <- c(
  ## R Markdown
  "kableExtra","DT","texreg",
  ## Tidyverse
  "tidyverse", "lubridate", "forcats", "haven", "labelled",
  ## Extensions for ggplot
  "ggmap","ggrepel", "ggridges", "ggthemes", "ggpubr", 
  "GGally", "scales", "dagitty", "ggdag", "ggforce",
  # Graphics:
  "scatterplot3d", #<<
  # Data 
  "COVID19","maps","mapdata","qss","tidycensus", "dataverse", 
  # Analysis
  "DeclareDesign", "easystats", "zoo"
)
```

```{r ipak, include=F}
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}

```


```{r loadpackages, cache=F, include=F}
ipak(the_packages)
```

---
class: inverse, center, middle
# Overview

---
## General Plan

- Setup
- Feedback
- Review
- Lecture 
  - Topic 1
  - Topic 2

- Summary


---
class:inverse, middle, center
# 💪
## Get set up to work

---
## New packages

Hopefully, you were all able to install the following packages 

```{r, eval=FALSE}


```



---
## Packages for today


```{r, ref.label=c("packages")}

```

---
## Define a function to load (and if needed install) packages


```{r, ref.label="ipak"}
```

---
## Load packages for today

```{r ref.label="loadpackages"}
```


---
class:inverse, center, middle
# 💪
## Load Data for today


# 📢
## Feedback


```{r feedback, echo=F, message=F}
df_feedback_full <- haven::read_spss("../files/data/wk05.sav")
df_feedback_full %>%
  filter(optout == 1)-> df_feedback
```


---
## What we liked

--

- We liked the lab (or liked it better)

- Working in groups 

- Extensive notes

- Reasonable pacing 

- Real world data

- Mix of substantive and technical questions

---
## What we liked

```{r likes, echo=F, out.height='90%'}
DT::datatable(df_feedback %>% 
                select(Likes),
               fillContainer = F,
              height = "90%",
              options = list(
                pageLength = 4
              )
              )
```


---
## What we disliked

--

- Labs still feel rushed (at the end)

- When code breaks it can be hard to fix

- Lecture was a bit confusing/long

- The space

---
## What we disliked

```{r dislikes, echo=F}
DT::datatable(df_feedback %>% 
                select(Dislikes),
               fillContainer = F,
              height = "90%",
              options = list(
                pageLength = 3
              )
              )
```


---
class:inverse, middle, center
# 🔍
## Review


---
## Topics

- Working with R Markdown files

- Basic Programming in R

- Setting up your work space

- Loading data into R

- Looking at your data

- Cleaning and transforming your data

- Describing what's typical of your data

- Describing how your data vary

---
## Exercises

1. Create an R Markdown File

2. Explore objects and functions in base `R`

3. Install the COVID19 package

4. Load data on COVID19 

5. Inspect the data

6. Welcome to the tidyverse

7. Measures of central tendency

8. Measures of dispersion

---
class: inverse, center, middle
# 💡
# Key Concept
## Comment


---
## Point 1


---
## Code Slide

.left-column[
```{r,echo=T,eval=F}
2+2
```


]

.right-column[
```{r,eval=T}
plot(1:10,1:10)
```

]

---
class: inverse, center, middle

# Key Concept
